---
title: 'Assignment 1: Data Preprocessing and Feature Engineering'
author: "Shagufta  Perveen"
date: "2025-04-155"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## 1. Dataset Selection

We use a **Customer Behavior dataset** from [Kaggle](https://www.kaggle.com/datasets). This dataset contains behavioral information about customers, which aligns with our research topic on customer segmentation and purchase prediction.

```{r load-data}
# Load the dataset
data <- read.csv("C:/Users/Buzz Tech/Downloads/Customer_Behaviour.csv")
head(data)
```

## 2. Exploratory Data Analysis (EDA)

### Descriptive Statistics

```{r summary-stats}
summary(data)
summary(data$Age)
```

### Visualizations

#### Histogram of Age

```{r histogram}
library(ggplot2)
ggplot(data, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  theme_minimal() +
  ggtitle("Age Distribution")
```

#### Boxplot of Age

```{r boxplot}
ggplot(data, aes(y = Age)) +
  geom_boxplot(fill = "lightblue") +
  theme_minimal() +
  ggtitle("Boxplot of Age")
```

#### Scatter Plot: Age vs Purchase

```{r scatterplot}
ggplot(data, aes(x = Age, y = Purchased)) +
  geom_point(color = "blue") +
  theme_minimal() +
  ggtitle("Age vs Purchase Amount")
```

### Variable Descriptions

```{r structure}
str(data)
```

## 3. Data Preprocessing

### Handling Missing Data

```{r missing-data}
# Remove rows with any missing values
data_clean <- na.omit(data)

# Impute missing Age values with median
data$Age[is.na(data$Age)] <- median(data$Age, na.rm = TRUE)
```

### Outlier Detection and Removal

```{r outlier-removal}
Q1 <- quantile(data$Age, 0.25)
Q3 <- quantile(data$Age, 0.75)
IQR_value <- Q3 - Q1

# Filter out outliers
data_clean <- data[data$Age >= (Q1 - 1.5 * IQR_value) & data$Age <= (Q3 + 1.5 * IQR_value), ]
```

### Data Transformation

```{r transformation}
# Standardization
data$Age_standardized <- scale(data$Age)

# Log transformation
data$Purchased_log <- log(data$Purchased + 1)
```

### Data Cleaning

```{r cleaning}
# Standardize text in 'Gender'
data$Gender <- tolower(data$Gender)
```

## 4. Feature Engineering

### Customer Lifetime Value (CLV)

```{r clv}
# Example: CLV = Frequency * Average Purchase Amount
data$Frequency * data$AvgPurchased
head(data$CLV)
```

### Recency Feature

```{r recency}
# Recency in days from LastPurchaseDate to today
 as.numeric(difftime(Sys.Date(), as.Date(data$LastPurchaseDate), units = "days"))
head(data$Recency)
```

## 5. Conclusion

We have successfully preprocessed the Customer Behavior dataset by handling missing values, detecting and treating outliers, performing transformations, and engineering new features like CLV and Recency. These steps will improve model performance and support deeper insights in our research analysis.
